Scheduling Musical Events as Procedures with Scheme for Max

### 2021-06-18 paper-1/paper-1.tex
wrote draft intro

\section{Introduction}
Here is the introduction about why Max and Scheme are so cool.
\section{Conclusion}
Well that was cool, I wonder how I should conclude such a thing???

>>>>

\section{Introduction}
Here is the introduction about why Max and Scheme are so cool.

Max (also known as Max/MSP) is a programming environment for
creating interactive music and multi-media programs through
a visual "patching" programming language, 
similar in appearance to a modular synthesizer
and accessible to non-programmer users. 
First created in the mid 1980's by Miller Puckette while at IRCAM,
it is now developed and sold by the San Francisco software company Cycling 74,
and is widely used in both academic and commercial music contexts as well
as in multi-media installations. %%citation?%%
A rich library of Max objects exists, both built in and available as open-source 
extensions, enabling users to rapidly create interactive systems with gestural 
input from physical sources such as MIDI controllers, custom hardward 
(over UDP networks), and on screen widgets.

%% TODO visual example of a Max patch with various inputs %%

One powerful feature of Max is its ability to be programmed
while the engine is playing music. Patches can be re-programmed in "edit-mode"
without necessarily interrupting patch activity (depending on the design
of the program), and this can
even be performed live, an activy known as "live-coding", which has given
birth to an entire musical sub-culture of live programming performances
with tools such as Max, SuperCollider, Csound, and others. 
%% citation%%
This approach to music making overlaps with the related discipline of 
"algorithmic music", in which programmatic algorithms are used
not just for affecting musical parameters such as volume and timber, 
but also for the generation of musical semantic content such as notes
and rhythms.

In both the fields of live-coding and algorithmic music, the ability for
the performer/composer to schedule events in the future with high temporal
accuracy is of major benefit. While this is possible in the Max visual
programming language, it is cumbersome and thus the implementation is not
optimal for live-coding or algorithmic music work where the speed with which
the user can accomplish this is an important consideration (i.e., fast
enough not to bore an audience or to be frustrating while composing).

In addition to the Max visual programming language, Max can be extended
by creating new Max objects (refered to as "externals") in C or C++ 
using the Max SDK and API, and can also be programmed in text-based 
languages through Max visual objects that embed language
interpreters. Max provides one such interpreter through its support for 
JavaScript in the 'js' object, and others contributed by 3rd parties
exist, such as for Lua, Python, and Ruby.

Scheme for Max (a.k.a S4M) is an open-source external developed by the 
authors that embeds the s7 Scheme interpreter, created by Bill Schottstaedt at
the CCRMA at Stanford University. Similar in functionality to the built
in js object, the s4m object enables the user to program Max in Scheme,
with facilities to interact with the Max engine and environment through
a foreign function interface implemented in C and calling the Max API functions.

We believe that s4m extends Max in a way that is highly valuable to 
the algorithmic composer and live-coding performer and is of great benefit
to the broader Max user community. The ability to update
a running Scheme program during playback is a major benefit
and is a capability has been, while technically possible, impractical
with pre-existing text language solutions such as JavaScript. The semantics
of Scheme bear a striking similarity to Max, and thus one can even use
Max visual widgets and messages to generate small Scheme programs in the
patcher. However, the most notable benefit is the ease with which the user 
can create and schedule functions for evaluation in the future, with 
input into these functions coming from Max patching widgets.

This paper provides an examination of the problem of musical scheduling
in computer music, and an overview of the Max platform and its solutions to this problem, 
along with a discussion of the limitations of these solutions.
It then introduces Scheme for Max and examines how the use of Scheme
overcomes these limitations elegantly, and provides details of
the implementations of scheduling in S4M, both at the Scheme and
C programming levels. Finally, it concludes with discussion of future
possibilities with Scheme for Max.
